---
title: "Lab"
author: "An Bui"
date: "2/14/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Goals: 

- Read in spatial data (shapefiles)
- Simplify polygons (st_simplify)
- Add and transform projection (st_transform and st_crs)
- Create several finalized-ish maps in R
- Join spatial data (st_join)
- Find spatial intersections (st_intersect)
- Interactive plots with tmap
- Updating color schemes, base layers, etc.
- Plotting spatial lines
- Creating sf data from lat/lon coordinates

```{r}
#### libraries ####
library(sf)
library(tidyverse)
library(tmap)
library(leaflet)
library(ggrepel)
library(ggspatial)
library(RColorBrewer)
library(raster)
```

##### Example 1. Dams in California

```{r get data into R}
#### California ecoregions ####
# lots of files that will need to be loaded, so plan accordingly
# read_sf() for simple features information (layers from files), "." for current working directory
ca_eco <- read_sf(".", 
                  # all files with ca_eco in them
                  layer = "ca_eco") %>% 
  # only select column with ecoregion name, sf automatically attaches geometries
  dplyr::select(US_L3NAME) %>% 
  # rename column to "Region"
  rename(Region = US_L3NAME) %>% 
  # simplify polygon to get files to run more easily
  st_simplify(dTolerance = 100) %>% 
  # specify coordinate reference system
  st_transform(crs = 4326)

# st_crs(ca_eco) gives summary info

#### California counties ####
ca_counties <- read_sf(".", layer = "california_county_shape_file")

# set same projection system as ca_eco
st_crs(ca_counties) = 4326

#### California dams ####
ca_dams <- read_sf(".", layer = "California_Jurisdictional_Dams") %>% 
  # rename column for ease of use
  rename(Condition = Condition_)

# set condition column to factor
# choose column
ca_dams$Condition <- fct_relevel(ca_dams$Condition, 
                                 # set levels
                                 "Fair", "Satisfactory", "Unsatisfactory", "Poor")
```

```{r maps}
#### plot in base R ####
plot(ca_eco)

plot(ca_counties)
# makes different maps for each attribute

#### map in ggplot ####
# modify RColorBrewer palettes
color_count <- 13
# create a palette using brewer.pal(number of original colors to use, palette)
new_colors <- colorRampPalette(brewer.pal(10, "Set2"))(color_count)

ca_eco_plot <- ggplot(ca_eco) +
  # create a map using spatial information
  geom_sf(aes(fill = Region),
          # get rid of outline
          color = "NA",
          # get rid of legend
          show.legend = FALSE) +
  # change color scheme
  scale_fill_manual(values = new_colors) +
  # add geom_sf() for counties
  geom_sf(data = ca_counties,
          # take out fill, only leave county lines
          fill = "NA",
          # change outline color
          color = "gray30",
          # change size of line
          size = 0.1) +
  # add dam locations
  geom_point(data = ca_dams,
             # specify aes because there is longitude and latitude data to map dams
             aes(x = Longitude, y = Latitude),
             size = 1,
             color = "gray10",
             alpha = 0.5) +
  theme_minimal() +
  # get rid of coordinates on axes
  coord_sf(datum = NA)
  
  
ca_eco_plot
```

##### Example 2. Dams in the Sierra Nevada eco-region
```{r}
#### wrangling ####
sn <- ca_eco %>% 
  filter(Region == "Sierra Nevada") %>% 
  # join with ca_dams information
  st_join(ca_dams)

#### plot ####
ggplot(sn) +
  # take all ca_counties data
  geom_sf(data = ca_counties, 
          fill = "wheat3",
          color = "NA") +
  # take sn data specifically
  geom_sf(fill = "lemonchiffon4",
          color = "NA") +
  # add dams
  geom_point(aes(x = Longitude, y = Latitude),
             size = 0.5,
             color = "red4")
```

##### Example 3. Eco-regions for Santa Barbara County

```{r eco-regions for SBCo}
#### wrangle for SBCo ####
sb <- ca_counties %>% 
  filter(NAME == "Santa Barbara")
# four rows because islands

#### clip eco-region data for regions only in SBCo ####
# clip ca_eco polygon by the bounds of the sb polygon
eco_clip <- st_intersection(ca_eco, sb)
plot(eco_clip)

#### more finalized plot ####
ggplot(eco_clip) +
  # county information
  geom_sf(data = ca_counties, 
          fill = "gray90",
          color = "gray80",
          size = 0.2) +
  # add clipped ecoregion data
  geom_sf(aes(fill = Region),
          color = "NA") +
  # change color scheme
  scale_fill_manual(values = c("darkolivegreen2", "darkolivegreen", "gold2")) +
  # limit coordinates in map
  coord_sf(xlim = c(-121, -119), 
           ylim = c(33.5, 35.5)) +
  # add point for downtown SB
  geom_point(aes(x = -119.6982, y = 34.4208),
             size = 2) +
  # add a label
  geom_text(x = -119.6982, y = 34.35,
            label = "Santa Barbara") +
  theme_minimal() +
  theme(legend.position = c(0.5, 0.15))
```

##### Example 4. Intro to interactive plots with `tmap`

```{r plot with tmap}
#### tmap ####
map_sb_eco <- tm_shape(eco_clip) +
  # fill polygons by Region
  tm_fill("Region", 
          palette = "RdPu", 
          alpha = 0.5) +
  # plot borders
  # for each layer, set a different shape
  tm_shape(ca_counties) +
  tm_borders()

# sets tmap_mode to interactive viewing
tmap_mode("view")

map_sb_eco

#### try with a different base map #####
# leaflet::providers gives a list of all possible basemaps
tm_basemap("CartoDB.DarkMatter") +
  tm_shape(eco_clip) +
  tm_borders(col = "white")
```

##### Example 5. Fault lines

```{r}
#### read in data ####
fault_lines <- read_sf(".", layer = "GMC_str_arc") %>% 
  # use if there is an existing projection (.prj file)
  st_transform(crs = 4326) %>% 
  # break LTYPE into separate columns
  separate(LTYPE, into = c("syn_ant", "certainty", "direction"),
           sep = ",")
```













